---
title: "HW2"
author: "Semyon Nikulin"
date: "2025-01-27"
output:
  html_document: default
  pdf_document: default
---

UT eid: syn228

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, echo=FALSE, message=FALSE}
library(ggplot2)
library(tidyverse)
library(knitr)
library(kableExtra)

profs <- read.csv("profs.csv")

ggplot(profs, aes(x = eval)) +
  geom_histogram(binwidth = 0.2, fill = "skyblue", color = "black") +
  labs(
    title = "Distribution of Course Evaluation Scores",
    x = "Evaluation Scores",
    y = "Frequency"
  ) +
  theme_minimal()
```

This histogram graphs the distribution of evaluation scores for a sample of UT professors. This distribution appears to be approximately normal (although there is a slight left skew), and is centered at a mean of `r round(mean(profs$eval),2)`.

```{r, echo=FALSE}
ggplot(profs, aes(x = native, y = eval, fill = native)) +
  geom_boxplot() +
  labs(
    title = "Distribution of Course Evaluation Scores by Native English Speaker",
    x = "Native English Speaker",
    y = "Evaluation Scores"
  ) +
  scale_fill_manual(values = c("skyblue", "orange")) +
  theme_minimal()
```

This boxplot graphs the distribution of evaluation scores grouped by whether the professor is a native english speaker or not. Native speakers appear to have higher evaluation scores with a median of 4, while nonnative speakers appear to have lower evaluation scores with a median of 3.7.

```{r, echo=FALSE}
ggplot(profs, aes(x = eval, fill = gender)) +
  geom_histogram(binwidth = 0.2, color = "black", alpha = 0.7) +
  facet_grid(gender ~ .) +
  labs(
    title = "Distribution of Course Evaluation Scores by Gender",
    x = "Evaluation Scores",
    y = "Frequency"
  ) +
  theme_minimal() +
  scale_fill_manual(values = c("skyblue", "orange"))
```

This faceted histogram graphs the distribution of evaluation scores of professors grouped by gender. Both plots have a slight left skew, and are both approximately centered around 4. 

```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(profs, aes(x = beauty, y = eval)) +
  geom_point(alpha = 0.7, color = "blue") +
  geom_smooth(method = "lm", se = TRUE, color = "red") +
  labs(
    title = "Scatterplot of Physical Attractiveness vs Course Evaluations",
    x = "Physical Attractiveness (Beauty)",
    y = "Course Evaluation Scores"
  ) +
  theme_minimal()
```

This scatterplot graphs the physical attractiveness of the professor against their course evaluation scores. There is a weak relationship between these two variables, with a correlation coefficient of `r round(cor(profs$beauty, profs$eval),2)`.

```{r echo=FALSE, message=FALSE, warning=FALSE}
bikes = read.csv("bikeshare.csv")
average_hourly_rentals <- bikes %>%
  group_by(hr) %>%
  summarize(average_rentals = mean(total))

ggplot(average_hourly_rentals, aes(x = hr, y = average_rentals)) +
  geom_line(color = "blue", size = 1) +
  labs(title = "Average Hourly Bike Rentals",
       x = "Hour of the Day",
       y = "Average Number of Rentals") +
  theme_minimal() +
  scale_x_continuous(breaks = 0:23) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

This line graph plots the average hourly bike rentals across all hours of the day. These bike rentals peak around 8 and 17 hours into the day. Looking at the graph, we can see that bike rentals are more popular during work transit hours (8 am and 5 pm) and are more popular during the day.

```{r echo=FALSE}
average_bike_rentals <- bikes %>%
  group_by(hr, workingday) %>%
  summarize(average_rentals = mean(total), .groups = "drop")

ggplot(average_bike_rentals, aes(x = hr, y = average_rentals)) +
  geom_line(color = "blue", size = 1) +
  facet_wrap(~workingday, labeller = labeller(workingday = c("0" = "Non-Workday", "1" = "Workday"))) +
  labs(
    title = "Average Bike Rentals by Hour of the Day",
    x = "Hour of the Day",
    y = "Average Bike Rentals"
  ) +
  theme_minimal() +
  scale_x_continuous(breaks = 0:23) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6))
```

This line graph also plots the average hourly bike rentals across all hours of the day, but is faceted according to whether it is a working day. Our conclusion from the last graph was correct, as the peaks around 8am and 5pm are shown in the working days graph, while hourly bike rentals peak during midday on non-workdays. As we can see, average bike patterns differ on workdays and non-workdays due to transportation patterns.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
filtered_data <- bikes %>%
  filter(hr == 9)

ggplot(filtered_data, aes(x = weathersit, y = total, fill = as.factor(workingday))) +
  stat_summary(fun = "mean", geom = "bar", position = "dodge") +
  scale_fill_manual(values = c("steelblue", "coral"), labels = c("Non-working Day", "Working Day")) +
  labs(
    title = "Average Ridership at 9 AM by Weather Situation",
    x = "Weather Situation",
    y = "Average Ridership",
    fill = "Working Day"
  ) +
  theme_minimal()
```

This bar plot shows the average ridership across different weather situations, faceted according to whether it was a workday or not. There were the most riders in weather situation 1 (clear weather) and there were the least riders in weather situation 3 (light snow/rain). The workday trend of higher traffic at 9am continues in this graph. Looking at this graph, we can see that there is more ridership in better weather.

\newpage
```{r echo=FALSE, message=FALSE, warning=FALSE}
data <- read.csv("capmetro_UT.csv")

average_boardings <- data %>%
  group_by(hour_of_day, day_of_week, month) %>%
  summarise(avg_boarding = mean(boarding, na.rm = TRUE)) %>%
  ungroup()

ggplot(average_boardings, aes(x = hour_of_day, y = avg_boarding, color = month)) +
  geom_line(size = 1) +
  facet_wrap(~ day_of_week, ncol = 3) +
  labs(
    title = "Average Boardings by Hour of Day, Faceted by Day of Week",
    x = "Hour of Day",
    y = "Average Boardings",
    color = "Month"
  ) +
  theme_minimal() +
  theme(
    strip.text = element_text(size = 12),
    axis.text = element_text(size = 10),
    axis.title = element_text(size = 12),
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 10)
  )
```

This figure displays the average boardings of Austin's Capital Metro bus network by hours of the day, grouped by month and faceted by day of the week. Looking at the graph, we can see that on weekdays boardings peak around 3pm (and there is no peak on weekends). Average boardings on Mondays in September tend to be lower, likely due to better weather conditions, leading people to use other forms of transportation like bikes. Wednesdays, Thursdays, and Fridays in November have lower boarding rates, likely due to holidays like Thanksgiving.

\newpage
```{r, echo=FALSE, message=FALSE}

ggplot(data, aes(x = temperature, y = boarding, color = weekend)) +
  geom_point(alpha = 0.5) +
  facet_wrap(~ hour_of_day, ncol = 4) +
  scale_color_manual(values = c("weekday" = "blue", "weekend" = "red")) +
  labs(
    title = "Boardings vs. Temperature by Hour of Day",
    x = "Temperature",
    y = "Boardings",
    color = "Day Type"
  ) +
  theme_minimal() +
  theme(
    strip.background = element_rect(fill = "lightgray"),
    strip.text = element_text(face = "bold"),
    legend.position = "right"
  )
```

This figure shows the temperature vs boardings in the same bus network, faceted by hour of the day and grouped by whether it was a weekday or weekend. When we hold hour of day and weekend status constant, temperature does not seem to have a major effect on the number of UT students riding the bus, as the scatter plots do not seem to have strong correlations.


```{r, echo=FALSE}
billboard <- read.csv("billboard.csv")
top_10_songs <- billboard %>%
  group_by(performer, song) %>%
  arrange(desc(weeks_on_chart)) %>%
  head(10)

top_10_songs <- subset(top_10_songs, select = c("performer", "song", "weeks_on_chart"))
kable(top_10_songs, col.names = c("Performer", "Song", "Count"), caption = "Top 10 Songs with Most Weeks on Billboard Hot 100 (1958-2025)") %>%
  kable_styling(bootstrap_options = "bordered")
```

This table shows the songs with the longest time spent on the Billboard Top 100, where more popular songs appear at the top of the table. Imagine Dragons seems to be the most popular performer.

```{r echo=FALSE}
filtered_data <- billboard %>%
  filter(year != 1958 & year != 2021)

diversity_data <- filtered_data %>%
  group_by(year) %>%
  summarise(unique_songs = n_distinct(song))

ggplot(diversity_data, aes(x = year, y = unique_songs)) +
  geom_line() +
  labs(
    title = "Musical Diversity Over the Years",
    x = "Year",
    y = "Number of Unique Songs"
  ) +
  theme_minimal()
```

This line graph shows the musical diversity in the Top 100 Billboard over the years, where musical diversity is calculated by the number of unique songs on the billboard that year. It seems that musical diversity was at a decline from 1965-2000, but has resurfaced from 2000-2020.

```{r echo=FALSE}
ten_week_hits <- billboard %>%
  filter(weeks_on_chart >= 10)

ten_week_hits_count <- ten_week_hits %>%
  group_by(performer) %>%
  summarise(ten_week_hits_count = n_distinct(song))

top_artists <- ten_week_hits_count %>%
  filter(ten_week_hits_count >= 30) %>%
  arrange(desc(ten_week_hits_count))

ggplot(top_artists, aes(x = reorder(performer, -ten_week_hits_count), y = ten_week_hits_count)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  coord_flip() +
  labs(
    title = "Top Artists with Ten-Week Hits",
    x = "Artist",
    y = "Number of Ten-Week Hits"
  ) +
  theme_minimal()
```

The figure shows the artists with the most songs on the Billboard 100 that stayed there for at least 10 weeks. The distribution of the amount of ten week hits seems approximately uniform across the top 19 artists. However, Elton John had the most ten-week hits.